---
title: "NetCDF_Attempt"
author: "James Waterford"
date: "2023-09-10"
output:
  pdf_document: default
  html_document: default
---

## NetCdf Installation

NetCdf is a file format widely used for acessing Earth Science Systems data. While the most common applications it is used with are **Java** and **C++**, Packages exist in both [R Studio](https://cirrus.ucsd.edu/~pierce/ncdf/index.html) and [Python](https://unidata.github.io/netcdf4-python/#groups-in-a-netcdf-file).

## R Install

To install in R, you will need both the `ncd4 package`, and the `NetCDF libraries`.

THIS STEP IS NOT REQUIRED AS `ncd4` INCLUDES ALL LIBRARIES

The NetCDF library can be found on the [UCAR](https://downloads.unidata.ucar.edu/netcdf/) website - Download the `netcdf-c-4.9.2.tar.gz` into a location you will remember - From the R-Studio dropdown menu, select: Tools \> Install Packages

```{r}
# install.packages("ncdf4")
library(ncdf4)
```

# File Connection

Use `nc_open` to read an existing netCDF file. nc_open(filename, write=FALSE)

For the dataset we are working with today, download from [USAP](https://www.usap-dc.org/view/dataset/601175). Additionally, the author of the USAP file suggests installing [Panoply](https://www.giss.nasa.gov/tools/panoply/download/) for the explanation of variables.

```{r openfile}

gen_atmos_file <- ("LENO2000H_10_yr_avg.nc")
gen_atmos <- nc_open(gen_atmos_file)
#print(gen_atmos)

gen_surf_file <- ("LENO2000H_LSX_10_yr_avg.nc")
gen_surf <- nc_open(gen_surf_file)
# print(gen_surf)
## Printing this line creates 6 pages in PDF format ##
```

In our case, we have 63 variables. Trying to determine all 63 named variables will prove to be a challenge. Use Panolopy to gain some insight.


Next Steps
-----------
** Create a Data Frame **

Now that we can access the data, we ned to extract the longitude and latitude values of interest from this full data set.
## Current Objectives: ##

 - Gather data temperature around the Drake passage 
 - There are twelve slices for each month
 - Extract temperatures over time
 -Calculate Mean Annual

Define Long & Lat
-----
```{r}
lon <- ncvar_get(gen_surf,"lon")
lon
lat <- ncvar_get(gen_surf,"lat")
lat
```



Create Temp Slice
------
```{r}
 ## Get Ocean Temperature from Surface Data
dname <- "TOCEAN"
tmp_array <- ncvar_get(gen_surf,dname)

# get a single slice or layer (January)
m <- 1
tmp_slice <- tmp_array[,,m]
```
## Map and Map Libraries

```{r}
library(lattice)
library(RColorBrewer)
```
This will map all temperature varation across the full dataset.
```{r}
grid <- expand.grid(lon=lon, lat=lat)
cutpts <- c(0,150,280,290,295,300,302.5,305)
levelplot(tmp_slice ~ lon * lat, data=grid, at=cutpts, cuts=9, pretty=T, 
  col.regions=(rev(brewer.pal(9,"YlOrBr"))))
```
Now we need to gather the relevant points into a data frame.
```{r}
# create dataframe -- reshape data


## Create Full data set from map & temperature
lonlat <- as.matrix(expand.grid(lon,lat))
tmp_vec <- as.vector(tmp_slice)

tmp_df01 <- data.frame(cbind(lonlat,tmp_vec))
names(tmp_df01) <- c("lon","lat",paste(dname,as.character(m), sep="_"))


## Subset to Drake Passage 
## [lon -> 240.00 : 300.00   | lat -> -89.00 : -15.00]

tmp_df01 <- tmp_df01[tmp_df01$lat <= -15 & tmp_df01$lon >= 240 & tmp_df01$lon <= 300,]
tmp_df01[tmp_df01 == 0] <- NA
head(na.omit(tmp_df01), 10)
```
Using functions to streamline this process.
------


```{r}

lonlat <- as.matrix(expand.grid(lon,lat))
tmp_df <- data.frame(cbind(lonlat))
names(tmp_df) <- c("lon","lat")

testr <- function(tmp_df){
  for (i in 1:12) {
    m <- i
    mon <- month.name[i]
    tmp_slice <- tmp_array[,,m]
    tmp_vec <- as.vector(tmp_slice)
    tmp_df[[mon]] <- tmp_vec
  }
  tmp_df ->> DP_tmp
}
testr(tmp_df)
DP_tmp <- DP_tmp[DP_tmp$lat <= -15 & DP_tmp$lon >= 240 & DP_tmp$lon <= 300,]
DP_tmp[DP_tmp == 0] <- NA

head(na.omit(DP_tmp))

write.csv(DP_tmp, "DrakePassageTemps.csv", row.names=FALSE)

```


