---
title: "NetCDF_Attempt"
author: "James Waterford"
date: "2023-09-10"
output: html_document
---

## NetCdf Installation

NetCdf is a file format widely used for acessing Earth Science Systems data. While the most common applications it is used with are **Java** and **C++**, Packages exist in both [R Studio](https://cirrus.ucsd.edu/~pierce/ncdf/index.html) and [Python](https://unidata.github.io/netcdf4-python/#groups-in-a-netcdf-file).

## R Install

To install in R, you will need both the `ncd4 package`, and the `NetCDF libraries`.

THIS STEP IS NOT REQUIRED AS `ncd4` INCLUDES ALL LIBRARIES

The NetCDF library can be found on the [UCAR](https://downloads.unidata.ucar.edu/netcdf/) website - Download the `netcdf-c-4.9.2.tar.gz` into a location you will remember - From the R-Studio dropdown menu, select: Tools \> Install Packages

```{r}
# install.packages("ncdf4")
```

# File Connection

Use `nc_open` to read an existing netCDF file. nc_open(filename, write=FALSE)

For the dataset we are working with today, download from [USAP](https://www.usap-dc.org/view/dataset/601175). Additionally, the author of the USAP file suggests installing [Panoply](https://www.giss.nasa.gov/tools/panoply/download/) for the explanation of variables.

```{r openfile}

gen_atmos_file <- ("LENO2000H_10_yr_avg.nc")
gen_atmos <- nc_open(gen_atmos_file)
#print(gen_atmos)

gen_surf_file <- ("LENO2000H_LSX_10_yr_avg.nc")
gen_surf <- nc_open(gen_surf_file)
print(gen_surf)

```

In our case, we have 63 variables. Trying to determine all 63 named variables will prove to be a challenge. Use Panolopy to gain some insight.


Next Steps
-----------

 - Gather data temperature around the Drake passage 
 - You have a period of twelve months.
 - Extract temperatures over time
 -Calculate Mean Annual


------

Grab values from specifiied long & lat.

Create twelve datasets, one for each month.

Interjoin long and lat over time to create seasonality.

```{r}
DPTO1 <- read.csv(file = "DrakePassageT_Ocean_Jan.csv", header=TRUE)
Lat <- colnames(DPTO1)
cnt <- nrow(DPTO1)
DPTJan <- data.frame()
DPTO1 %>%
  lt <- 
  DPTJan$lat <- as.string(Lat[i][2:])
  DPTJan$long <- X[i]
DPTJan$Jan <- DPTO1$
  for line in 
```



